### 2014-04-27 stat246/hw3.tex
s3

\[\sum_{i=1}^N\sum_{m=1}^dw_{i,m}\left(\log\pi_m+\sum_{j=1}^dx_{\rho^{-m}(j)}\log\theta_j-\theta_j-\log(x_{\rho^{-m}(j)})\right)\]
Differentiating wrt $\pi_m$ and setting to zero gives $\h{\pi}_m=\frac{1}{N}\sum_{i=1}^Nw_{i,m}$. Differentiating wrt $\theta_j$ produces
\[\sum_{i=1}^N\sum_{m=1}^d\left(w_{i,m}\frac{x_{\rho^{-m}(j)}}{\theta_j}-1\right)=\theta_j\sum_{i=1}^N\sum_{m=1}^dw_{i,m}x_{\rho^{-m}(j)}-Nd\]
so setting this to zero produces $\theta_j=\frac{Nd}{\sum_{i=1}^N\sum_{m=1}^dw_{i,m}x_{\rho^{-m}(j)}}$
%TODO: computation
\subsection*{3}
\ssn{a}
We have 
\[\kl(X\|U)=\sum_{i=1}^KP(X=k)\log\left(\frac{K^{-1}}{P(X=k)}\right)=-\sum_{i=1}^KP(X=k)\log(KP(X=k))=-\sum_{i=1}^KP(X=k)\log P(X=k)-\log K\]
so $\log K-\kl(X\|U)=\sum_{i=1}^KP(X=k)\log P(X=k)=H(X)$.

>>>>

\[\sum_{i=1}^N\sum_{m=1}^dw_{i,m}\left(\log\pi_m+\sum_{j=1}^dx_{\rho^{-m}(j)}\log\theta_j-\theta_j-\log(x_{\rho^{-m}(j)})\right)\]
Differentiating wrt $\pi_m$ and setting to zero gives $\h{\pi}_m=\frac{1}{N}\sum_{i=1}^Nw_{i,m}$. Differentiating wrt $\theta_j$ produces
\[\sum_{i=1}^N\sum_{m=1}^d\left(w_{i,m}\frac{x_{\rho^{-m}(j)}}{\theta_j}-w_{i,m}\right)=\theta_j^{-1}\sum_{i=1}^N\sum_{m=1}^dw_{i,m}x_{\rho^{-m}(j)}-N\]
so setting this to zero produces $\theta_j=N^{-1}\sum_{i=1}^N\sum_{m=1}^dw_{i,m}x_{\rho^{-m}(j)}$, where there's an implicit $i$ subscript on the $x$

Below are index plots of $\theta$ and $\pi$. The first row is for $N=10$, descending to $N=20$ and $N=200$, and the true values are at the bottom. As might be expected, some of the results differ from the true value by a cyclic permutation, since the model is invariant under such permutations. The $\theta$s are on the left, the $\pi$s on the right.

\includegraphics[width=0.5\textwidth]{hw3_files/theta_10.png}
\includegraphics[width=0.5\textwidth]{hw3_files/pi_10.png}

\includegraphics[width=0.5\textwidth]{hw3_files/theta_20.png}
\includegraphics[width=0.5\textwidth]{hw3_files/pi_20.png}

\includegraphics[width=0.5\textwidth]{hw3_files/theta_200.png}
\includegraphics[width=0.5\textwidth]{hw3_files/pi_200.png}

\includegraphics[width=0.5\textwidth]{hw3_files/true_theta.png}
\includegraphics[width=0.5\textwidth]{hw3_files/true_pi.png}

\subsection*{3}
\ssn{a}
We have 
\[\kl(X\|U)=\sum_{i=1}^KP(X=k)\log\left(\frac{K^{-1}}{P(X=k)}\right)=-\sum_{i=1}^KP(X=k)\log(KP(X=k))=-\sum_{i=1}^KP(X=k)\log P(X=k)-\log K\]
so $\log K-\kl(X\|U)=\sum_{i=1}^KP(X=k)\log P(X=k)=H(X)$.
